Preparing your manuscript

### 2023-09-29 paper/paper.qmd
Update paper

    - Be set as a single paragraph.
    - Be less than 250 words for all journals except GRL, for which the limit is 150 words.
    - Not include table or figure mentions.
    - Avoid reference citations unless dependent on or directly related to another paper (e.g., companion, comment, reply, or commentary on another paper(s)). AGU’s Style Guide discusses formatting citations in abstracts.
    - Define all abbreviations.
plain-language-summary: |
  A Plain Language Summary (PLS) can be an incredibly effective science communication tool. By summarizing your paper in non-technical terms, you can explain your research and its relevance to a much broader audience. A PLS is required for submissions to AGU Advances, G-Cubed, GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of Geophysics, but optional for other journals.  A PLS should be no longer than 200 words and should be free of jargon, acronyms, equations, and any technical information that would be unknown to people from outside your scientific discipline. Read our tips for creating an effective PLS.
keywords: []
key-points:
  - Key Points convey the main points and conclusions of the article. 
  - Up to three key point statements are allowed, and each is limited to at most 140 characters with no abbreviations.
  - Key Points are not included in the word count.
bibliography: bibliography.bib  
citation:
  container-title: Geophysical Research Letters
keep-tex: true
date: last-modified
---
## Introduction

@@

The Long Short-Term Memory (LSTM), a specialized form of Recurrent Neural Networks (RNNs), is specifically tailored for modeling temporal sequences. Its unique design allows it to adeptly handle long-range dependencies, setting it apart from traditional RNNs in terms of accuracy. This prowess in modeling long-range dependencies has been validated in various studies [12, 11, 17, 23]. The cornerstone of LSTM's innovation is its memory cell, $c_t$t, which functions as a repository of state information, also refered to as long-term memory. This cell is accessed, modified, and reset through several self-parameterized gates. As new inputs arrive, their information is accumulated into the cell if the input gate is activated. The forget gate $f_t$ defines the percentage of the previous cell status that is stored $c_{t-1}$​. The decision to propagate the latest cell output, $c_t$, to the final state, $h_t$, is governed by the output gate, $o_t$. A significant advantage of this architecture is the memory cell's ability to retain gradients. This mechanism addresses the vanishing gradient problem, where, as input sequences elongate, the influence of initial stages becomes harder to capture, causing gradients of early input points to approach zero. The LSTM's activation function, inherently recurrent, mirrors the identity function with a consistent derivative of 1.0, ensuring the gradient remains stable throughout backpropagation.

@@

@@

The FC-LSTM, while adept in many scenarios, falters in encoding spatial information when handling spatiotemporal data due to its reliance on full connections in both input-to-state and state-to-state transitions. Addressing this limitation, our design ensures that all inputs $X_1, \ldots, X_t$, cell outputs $C_1, \ldots, C_t$, hidden states $H_1, \ldots, H_t$, and gates $i_t, f_t, o_t$ of the ConvLSTM are 3D tensors. The last two dimensions of these tensors represent spatial dimensions, specifically rows and columns. Conceptually, these inputs and states can be visualized as vectors positioned on a spatial grid.

@@

Considering the states as representations of moving entities, a ConvLSTM with a larger transitional kernel can capture rapid movements, while a smaller kernel is apt for slower motions. Drawing parallels with [16], the traditional FC-LSTM's inputs, cell outputs, and hidden states can also be envisioned as 3D tensors, with the last two dimensions being unitary. Thus, the FC-LSTM can be seen as a specific instance of ConvLSTM where all features are concentrated on a singular cell.

To maintain consistency in the dimensions of states and inputs, padding is essential before convolution operations. This padding, especially at boundary points, can be interpreted as leveraging the state of external surroundings for computations. Typically, all LSTM states are initialized to zero before the first input, symbolizing a lack of knowledge about future events. Similarly, zero-padding (as adopted in this study) on hidden states sets the external world's state to zero, implying no prior information about external conditions. This padding approach allows for differential treatment of boundary points, proving advantageous in many scenarios. For instance, in a system with a ball moving and bouncing off unseen walls, the existence of these walls can be inferred by observing the ball's repeated bounces, a deduction challenging if boundary points share the same state transition dynamics as inner points.

>>>>

plain-language-summary: |
  A Plain Language Summary (PLS) can be an incredibly effective science communication tool. By summarizing your paper in non-technical terms, you can explain your research and its relevance to a much broader audience. A PLS is required for submissions to AGU Advances, G-Cubed, GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets, JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of Geophysics, but optional for other journals.  A PLS should be no longer than 200 words and should be free of jargon, acronyms, equations, and any technical information that would be unknown to people from outside your scientific discipline. Read our tips for creating an effective PLS.
keywords: []
key-points:
  - Key Points convey the main points and conclusions of the article. 
  - Up to three key point statements are allowed, and each is limited to at most 140 characters with no abbreviations.
  - Key Points are not included in the word count.
bibliography: bibliography.bib  
citation:
  container-title: Geophysical Research Letters
keep-tex: true
date: last-modified
---
## Introduction

@@

The Long Short-Term Memory (LSTM), a specialized form of Recurrent Neural Networks (RNNs), is specifically tailored for modeling temporal sequences. Its unique design allows it to adeptly handle long-range dependencies, setting it apart from traditional RNNs in terms of accuracy (see @fig-lstm).

![Inner structure of a Long Short-Term Memory Cell](draw_lstm.pdf){#fig-lstm}

This performance in modeling long-range dependencies has been validated in various studies. The key component of LSTM's innovation is its memory cell, $c_t$t, which stores state information, also refered to as long-term memory. This cell is accessed, modified, and reset through several self-parameterized gates. For the input of the sequence  $x_t$ input, the forget gate $f_t$ defines the percentage of the previous long-term memory status $c_{t-1}$ that should be retained stored​. Next the input gate $i_t$ decides how much of the input is added to the the long-term memory, forming the updated cell state $c_{t}$. The decision to propagate the latest cell output, $c_t$, to the final state, $h_t$, is governed by the output gate, $o_t$, representing the updated short-term memory of the hidden state $h_t$. A significant advantage of this architecture is the memory cell's ability to retain gradients. This mechanism addresses the vanishing gradient problem, where, as input sequences elongate, the influence of initial stages becomes harder to capture, causing gradients of early input points to approach zero. The LSTM's activation function, inherently recurrent, mirrors the identity function with a consistent derivative of 1.0, ensuring the gradient remains stable throughout backpropagation.

@@

with

- $x_t$: Input vector at time step $t$.
- $h_{t-1}$: Hidden state from the previous time step.
- $C_{t-1}$: Cell state from the previous time step.
- $W$ and $b$: Weight matrices and bias vectors, respectively, associated with the gates of the LSTM. The subscripts denote the specific gate or operation they are associated with (e.g., $W_f$ and $b_f$ are the weight matrix and bias for the forget gate, respectively).
- $\sigma$: Sigmoid activation function, which squashes values between 0 and 1.
- $\tanh$: Hyperbolic tangent activation function, which squashes values between -1 and 1

@@

The FC-LSTM fails to handle information when handling spatiotemporal data due to its reliance on full connections in both input-to-state and state-to-state transitions. To adress this limitation we use a convLSTM architecture. convLSTM replaces the fully connected operations in the LSTM with convolutional operations. Hence, all inputs $X_1, \ldots, X_t$, cell outputs $C_1, \ldots, C_t$, hidden states $H_1, \ldots, H_t$, and gates $i_t, f_t, o_t$ of the ConvLSTM are 3D tensors. The last two dimensions of these tensors represent spatial dimensions, specifically rows and columns. Conceptually, these inputs and states can be visualized as vectors positioned on a spatial grid.

@@

In summary, the ConvLSTM excels at processing tasks that demand a combined understanding of spatial patterns and temporal sequences in data. It merges the image-processing capabilities of Convolutional Neural Networks (CNNs) with the time-series modeling of Long Short-Term Memory (LSTM) networks.

### Implemented model architecture

We first implemented the ConvLSTM using and encoder/decoder structure as discussed in ... . To predict all 97 rivers entering the Baltic Sea at once, we flatten the output and use fully connected layers resulting in 97 outputs.

For the computation we use the following set of hyper parameters:

| Parameter name | Parameter size | 
|------|------|
| Num. timesteps     | 32    |
| Conv. Kernelsize    | (7,7)    |
| Num. ConvLSTM layers    | 4    |
| Batch size| 32 |
| Learning Rate | 1e-4 with CosineAnnealing |

: Hyperparameters {#tbl-letters}

### 2023-09-29 paper/paper.tex
Update paper

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{agujournal2019}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{lineno}
\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
\usepackage{soul}
\linenumbers
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Preparing your manuscript},
  pdfauthor={Florian Börgel; Sven Karsten},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\journalname{Geophysical Research Letters}

\draftfalse

\begin{document}
\title{Preparing your manuscript}

\authors{Florian Börgel\affil{1}, Sven Karsten\affil{1}}
\affiliation{1}{Leibniz-Institute for Baltic Sea Research Warnemünde, }
\correspondingauthor{Florian Börgel}{florian.boergel@io-warnemuende.de}

\begin{abstract}
The abstract (1) states the nature of the investigation and (2)
summarizes the important conclusions. The abstract should be suitable
for indexing. Your abstract should:

\begin{itemize}
\tightlist
\item
  Be set as a single paragraph.
\item
  Be less than 250 words for all journals except GRL, for which the
  limit is 150 words.
\item
  Not include table or figure mentions.
\item
  Avoid reference citations unless dependent on or directly related to
  another paper (e.g., companion, comment, reply, or commentary on
  another paper(s)). AGU's Style Guide discusses formatting citations in
  abstracts.
\item
  Define all abbreviations.
\end{itemize}
\end{abstract}

\section*{Plain Language Summary}
A Plain Language Summary (PLS) can be an incredibly effective science
communication tool. By summarizing your paper in non-technical terms,
you can explain your research and its relevance to a much broader
audience. A PLS is required for submissions to AGU Advances, G-Cubed,
GeoHealth, GRL, JAMES, JGR: Biogeosciences, JGR: Oceans, JGR: Planets,
JGR: Solid Earth, JGR: Atmospheres, Space Weather, and Reviews of
Geophysics, but optional for other journals. A PLS should be no longer
than 200 words and should be free of jargon, acronyms, equations, and
any technical information that would be unknown to people from outside
your scientific discipline. Read our tips for creating an effective PLS.

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

\hypertarget{lstm-network}{%
\subsection{LSTM network}\label{lstm-network}}

The Long Short-Term Memory (LSTM), a specialized form of Recurrent
Neural Networks (RNNs), is specifically tailored for modeling temporal
sequences. Its unique design allows it to adeptly handle long-range
dependencies, setting it apart from traditional RNNs in terms of
accuracy. This prowess in modeling long-range dependencies has been
validated in various studies {[}12, 11, 17, 23{]}. The cornerstone of
LSTM's innovation is its memory cell, \(c_t\)t, which functions as a
repository of state information, also refered to as long-term memory.
This cell is accessed, modified, and reset through several
self-parameterized gates. As new inputs arrive, their information is
accumulated into the cell if the input gate is activated. The forget
gate \(f_t\) defines the percentage of the previous cell status that is
stored \(c_{t-1}\)\hspace{0pt}. The decision to propagate the latest
cell output, \(c_t\), to the final state, \(h_t\), is governed by the
output gate, \(o_t\). A significant advantage of this architecture is
the memory cell's ability to retain gradients. This mechanism addresses
the vanishing gradient problem, where, as input sequences elongate, the
influence of initial stages becomes harder to capture, causing gradients
of early input points to approach zero. The LSTM's activation function,
inherently recurrent, mirrors the identity function with a consistent
derivative of 1.0, ensuring the gradient remains stable throughout
backpropagation.

@@

\hypertarget{convlstm-network}{%
\subsection{ConvLSTM network}\label{convlstm-network}}

The FC-LSTM, while adept in many scenarios, falters in encoding spatial
information when handling spatiotemporal data due to its reliance on
full connections in both input-to-state and state-to-state transitions.
Addressing this limitation, our design ensures that all inputs
\(X_1, \ldots, X_t\), cell outputs \(C_1, \ldots, C_t\), hidden states
\(H_1, \ldots, H_t\), and gates \(i_t, f_t, o_t\) of the ConvLSTM are 3D
tensors. The last two dimensions of these tensors represent spatial
dimensions, specifically rows and columns. Conceptually, these inputs
and states can be visualized as vectors positioned on a spatial grid.

@@

Considering the states as representations of moving entities, a ConvLSTM
with a larger transitional kernel can capture rapid movements, while a
smaller kernel is apt for slower motions. Drawing parallels with
{[}16{]}, the traditional FC-LSTM's inputs, cell outputs, and hidden
states can also be envisioned as 3D tensors, with the last two
dimensions being unitary. Thus, the FC-LSTM can be seen as a specific
instance of ConvLSTM where all features are concentrated on a singular
cell.

To maintain consistency in the dimensions of states and inputs, padding
is essential before convolution operations. This padding, especially at
boundary points, can be interpreted as leveraging the state of external
surroundings for computations. Typically, all LSTM states are
initialized to zero before the first input, symbolizing a lack of
knowledge about future events. Similarly, zero-padding (as adopted in
this study) on hidden states sets the external world's state to zero,
implying no prior information about external conditions. This padding
approach allows for differential treatment of boundary points, proving
advantageous in many scenarios. For instance, in a system with a ball
moving and bouncing off unseen walls, the existence of these walls can
be inferred by observing the ball's repeated bounces, a deduction
challenging if boundary points share the same state transition dynamics
as inner points.

\hypertarget{acknowledgments}{%
\section{Acknowledgments}\label{acknowledgments}}

@@

\hypertarget{open-research}{%
\section{Open research}\label{open-research}}

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\vspace{1em}

\end{CSLReferences}

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\KOMAoption{captions}{tableheading}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Preparing your manuscript},
  pdfauthor={Florian Börgel; Sven Karsten},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{Preparing your manuscript}
\author{Florian Börgel \and Sven Karsten}
\date{9/29/23}

\begin{document}
\maketitle
\begin{abstract}
The abstract (1) states the nature of the investigation and (2)
summarizes the important conclusions. The abstract should be suitable
for indexing. Your abstract should:
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, enhanced, frame hidden, borderline west={3pt}{0pt}{shadecolor}, interior hidden, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

@@

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{lstm-network}{%
\subsubsection{LSTM network}\label{lstm-network}}

The Long Short-Term Memory (LSTM), a specialized form of Recurrent
Neural Networks (RNNs), is specifically tailored for modeling temporal
sequences. Its unique design allows it to adeptly handle long-range
dependencies, setting it apart from traditional RNNs in terms of
accuracy (see Figure~\ref{fig-lstm}).

\begin{figure}

{\centering \includegraphics{draw_lstm.pdf}

}

\caption{\label{fig-lstm}Inner structure of a Long Short-Term Memory
Cell}

\end{figure}

This performance in modeling long-range dependencies has been validated
in various studies. The key component of LSTM's innovation is its memory
cell, \(c_t\)t, which stores state information, also refered to as
long-term memory. This cell is accessed, modified, and reset through
several self-parameterized gates. For the input of the sequence \(x_t\)
input, the forget gate \(f_t\) defines the percentage of the previous
long-term memory status \(c_{t-1}\) that should be retained
stored\hspace{0pt}. Next the input gate \(i_t\) decides how much of the
input is added to the the long-term memory, forming the updated cell
state \(c_{t}\). The decision to propagate the latest cell output,
\(c_t\), to the final state, \(h_t\), is governed by the output gate,
\(o_t\), representing the updated short-term memory of the hidden state
\(h_t\). A significant advantage of this architecture is the memory
cell's ability to retain gradients. This mechanism addresses the
vanishing gradient problem, where, as input sequences elongate, the
influence of initial stages becomes harder to capture, causing gradients
of early input points to approach zero. The LSTM's activation function,
inherently recurrent, mirrors the identity function with a consistent
derivative of 1.0, ensuring the gradient remains stable throughout
backpropagation.

@@

with

\begin{itemize}
\tightlist
\item
  \(x_t\): Input vector at time step \(t\).
\item
  \(h_{t-1}\): Hidden state from the previous time step.
\item
  \(C_{t-1}\): Cell state from the previous time step.
\item
  \(W\) and \(b\): Weight matrices and bias vectors, respectively,
  associated with the gates of the LSTM. The subscripts denote the
  specific gate or operation they are associated with (e.g., \(W_f\) and
  \(b_f\) are the weight matrix and bias for the forget gate,
  respectively).
\item
  \(\sigma\): Sigmoid activation function, which squashes values between
  0 and 1.
\item
  \(\tanh\): Hyperbolic tangent activation function, which squashes
  values between -1 and 1
\end{itemize}

\hypertarget{convlstm-network}{%
\subsubsection{ConvLSTM network}\label{convlstm-network}}

The FC-LSTM fails to handle information when handling spatiotemporal
data due to its reliance on full connections in both input-to-state and
state-to-state transitions. To adress this limitation we use a convLSTM
architecture. convLSTM replaces the fully connected operations in the
LSTM with convolutional operations. Hence, all inputs
\(X_1, \ldots, X_t\), cell outputs \(C_1, \ldots, C_t\), hidden states
\(H_1, \ldots, H_t\), and gates \(i_t, f_t, o_t\) of the ConvLSTM are 3D
tensors. The last two dimensions of these tensors represent spatial
dimensions, specifically rows and columns. Conceptually, these inputs
and states can be visualized as vectors positioned on a spatial grid.

@@

In summary, the ConvLSTM excels at processing tasks that demand a
combined understanding of spatial patterns and temporal sequences in
data. It merges the image-processing capabilities of Convolutional
Neural Networks (CNNs) with the time-series modeling of Long Short-Term
Memory (LSTM) networks.

\hypertarget{implemented-model-architecture}{%
\subsubsection{Implemented model
architecture}\label{implemented-model-architecture}}

We first implemented the ConvLSTM using and encoder/decoder structure as
discussed in \ldots{} . To predict all 97 rivers entering the Baltic Sea
at once, we flatten the output and use fully connected layers resulting
in 97 outputs.

For the computation we use the following set of hyper parameters:

\hypertarget{tbl-letters}{}
\begin{longtable}[]{@{}ll@{}}
\caption{\label{tbl-letters}Hyperparameters}\tabularnewline
\toprule()
Parameter name & Parameter size \\
\midrule()
\endfirsthead
\toprule()
Parameter name & Parameter size \\
\midrule()
\endhead
Num. timesteps & 32 \\
Conv. Kernelsize & (7,7) \\
Num. ConvLSTM layers & 4 \\
Batch size & 32 \\
~Learning Rate & 1e-4 with CosineAnnealing \\
\bottomrule()
\end{longtable}

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

@@

\hypertarget{open-research}{%
\subsection{Open research}\label{open-research}}

@@

\hypertarget{references}{%
\subsection*{References}\label{references}}
\addcontentsline{toc}{subsection}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{0}{0}
\end{CSLReferences}
